---
title: "Variance partition of life span with different models"
author: "Fabio Morgante"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r set options, message=FALSE}
###Load libraries
library(ggplot2)
library(cowplot)
library(tidyr)
library(dplyr)

prefix <- "dgrp_lifespan_gxe"
```

## Introduction

The goal of this analysis is to assess the performance of the linear mixed models including genotype only, environment only,
genotype + environment, genotype + environment + genotype $\times$ environment information at predicting life span in the 
DGRP.

We used data from [Huang et al. (2020)](https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.3000645), which
has life span measurements for $n=176$ lines at 3 temperatures (18C, 25C, 28C) for the 2 sexes (coded as 1=females and 0=males), 
which results in $q=1,056$ observations. 
The authors showed that there is extensive $G \times E$ affecting life span. In the present work, we used both temperature 
and sex as $c=2$ environmental variables, for a total of $r=6$ environments.

We fitted the following linear mixed models:

* G-BLUP -- $y = \mu + g + \epsilon$.
* E-BLUP -- $y = \mu + e + \epsilon$.
* GE-BLUP -- $y = \mu + g + e + \epsilon$.
* GxE-BLUP -- $y = \mu + g + e + ge + \epsilon$.

where

$y$ is a $q$-vector of phenotypic observations, 
$\mu$ is a $q$-vector of intercept values, 
$g$ is a $q$-vector of additive genetic values, $g \sim N_q(\mathbf O, \mathbf{ZGZ^\intercal} \sigma^2_g)$, 
$\mathbf Z$ is a $q \times n$ incidence matrix, 
$\mathbf G$ is a $n \times n$ genomic relationship matrix (GRM), 
$e$ is a $q$-vector of environmental values, $e \sim N_q(\mathbf O, \mathbf{E} \sigma^2_e)$, 
$\mathbf E$ is a $q \times q$  matrix of similarity based on environmental variables, computed as 
$\mathbf E = \mathbf{XX}^\intercal$, 
$\mathbf X$ is a $q \times c$ matrix of environmental measurements,
$ge \sim N_q(\mathbf O, \mathbf{ZGZ^\intercal \circ E} \sigma^2_{ge})$,
$\epsilon$ is a $q$-vector of residual values, $\epsilon \sim N_q(\mathbf O, \mathbf{I} \sigma^2_\epsilon)$.

These models were fitted to the whole data using a Bayesian approach as implemented in the $\sf BGLR$ package.

## Results

```{r varcomps, fig.height=10, fig.width=13}
model <- c("G", "E", "GandE", "GxE")
pheno <- "original_pheno"

i <- 0

res_var <- as.data.frame(matrix(NA, ncol=6, nrow=4))
colnames(res_var) <- c("model", "VarG", "VarE", "VarGxE", "VarEps", "VarTot")


for(met in model){
  dat <- readRDS(paste0("output/", met, "_fit/", prefix, "_", pheno, "_", met, "_fit_whole_data.rds"))
  i <- i + 1

  res_var[i, 1] <- met
  if(!is.null(dat$ETA$G$varU)){
    res_var[i, 2] <- dat$ETA$G$varU
  } else {
    res_var[i, 2] <- NA
  }
      
  if(!is.null(dat$ETA$E$varU)){
    res_var[i, 3] <- dat$ETA$E$varU
  } else {
    res_var[i, 3] <- NA
  }

  if(!is.null(dat$ETA$GxE$varU)){
    res_var[i, 4] <- dat$ETA$GxE$varU
  } else {
    res_var[i, 4] <- NA
  }

  res_var[i, 5] <- dat$varE
  res_var[i, 6] <- sum(res_var[i, 2:5], na.rm=TRUE)
}


###PVE
res_var$propG <- res_var$VarG/res_var$VarTot
res_var$propE <- res_var$VarE/res_var$VarTot
res_var$propGxE <- res_var$VarGxE/res_var$VarTot
res_var$propEps <- res_var$VarEps/res_var$VarTot

res_var_long <- res_var %>% select(model, VarG, VarE, VarGxE, VarEps) %>% 
  gather(value="Var", key="Source", VarG, VarE, VarGxE, VarEps)

res_var_long <- transform(res_var_long, model=factor(model, levels=c("G", "E", "GandE", "GxE")),
                 Source=factor(Source, levels=c("VarEps", "VarGxE", "VarE", "VarG"), 
                               labels=c("VarRes", "VarGxE", "VarE", "VarG")))



p <- ggplot(res_var_long, aes(x = model, y = Var, fill = Source)) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_manual(values = c("lightgreen", "orange", "pink", "blue")) +
  labs(x = "Model", y = expression(italic(PVE)), fill="Source", title="") +
  theme_cowplot(font_size = 18) 

print(p)
```

